A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period

### 2021-05-06 analysis/paper/paper.Rmd
edit intro section based on R2's comments

<!--Our burial network and research implication-->
Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011]. Here we treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e. individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013]. Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.

@@

<!-- trade dynamic during this period and future work for comparison -->
We also recognize that trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have had contributed to social changes at Kiwulan. However, the current changes observed from burial networks and spatial distribution of trade items in the residential area at Kiwulan demonstrate consistent changes before and after the European arrival. Historical records also support the interaction and trade activities between Europeans and local Indigenous people. The trade items, including Anping jars, stonewares, Chinese porcelains, and beads found at Kiwulan were frequently found from archaeological sites with Dutch occupation, such as Fort Zeelandia. As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides an ideal example to explore the reaction of Indigenous people to European colonial presence at a settlement scale. We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Helping Dao and Tamsui. However, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence are rare to offer good comparison currently and future work is still needed.

>>>>

<!--Our burial network and research implication-->
Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011]. Here we use network analysis to explore similarities and differences between burials since it provides a statistical way for identifying underlying variable structuring burial relations. We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e. individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013]. Instead of exchange or transmission, our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg. @Collar2015, pp. 4]. This is because burial goods, especially high value ones, normally convey cultural symbol of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013]. Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.

@@

<!-- trade dynamic during this period and future work for comparison -->
We also recognize that trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have had contributed to social changes at Kiwulan. However, the current changes observed from burial networks and spatial distribution of trade items in the residential area at Kiwulan demonstrate consistent changes before and after the European arrival. Historical records also support the interaction and trade activities between Europeans and local Indigenous people. The trade items, including Anping jars, stonewares, Chinese porcelains, and beads found at Kiwulan were frequently found from archaeological sites with Dutch occupation, such as Fort Zeelandia. As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides an ideal example to explore the reaction of Indigenous people to European colonial presence at a settlement scale. We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Heping Dao and Tamsui. However, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence are rare to offer good comparison currently and future work is still needed.
